A boundary-integral framework to simulate \\ viscous erosion of a porous medium

### 2018-04-09 docs/methods/erosion_methods.tex
NM

At the first stage of RK2 (the half step) we discretize~\eqref{eq:mu}--\eqref{eqn:thetaevo3} as
\begin{align}
\label{eq:mu12d}
& \mu_k^{(n+1/2)} = \mu_k^{(n)} \exp \left( \eps k^2 \frac{\Dt}{2}
\zeta^{(n)} \right) \, , \\
\label{eq:L12d}
& \frac{L^{(n+1/2)} - L^{(n)}}{\Dt/2} = \Mterm^{(n)} \, , \\
\label{eq:th12d}
& \frac{ \mu_k^{(n+1/2)} \thhat_k^{(n+1/2)} - \mu_k^{(n)} \thhat_k^{(n)}}{\Dt/2} 
= \mu_k^{(n)} \hat{\NLterm}_k^{(n)} \, ,
\end{align}
where the superscript indicates the time step. At the next stage of RK2
it will be necessary to evaluate derivatives at the half step $n+1/2$, thus requiring a call to the Stokes solver for which the body shapes must be known. We therefore insert~\eqref{eq:mu12d} into~\eqref{eq:th12d} and solve for $\theta^{(n+1/2)}$, giving
\todo[inline]{are we solving for $\theta$ or $\hat{\theta}_k?$.}
\begin{align}
\label{eq:th12}
\thhat_k^{(n+1/2)} = \left( \thhat_k^{(n)} + \frac{1}{2} \Dt \, \hat{\NLterm}_k^{(n)} \right)
\exp \left( -\epsilon k^2 \frac{\Dt}{2} \elfun^{(n)}\right) \, .
\end{align}
Since $\zeta >0$, the exponential term always {\em suppresses} high-frequencies (exponentially with $k^2$). In other words, this is a smoothing term. With the geometry at step $n+1/2$ known, a call to the Stokes solver allows the computation of $\Mterm^{(n+1/2)}$, $\widehat{\NLterm}_k^{(n+1/2)}$, and $\zeta^{(n+1/2)}$.

At the second stage of RK2, we discretize~\eqref{eq:mu}--\eqref{eqn:thetaevo3} as 
\begin{align}
\label{eq:mu1d}
& \mu_k^{(n+1)} = \mu_k^{(n)} \exp \left( \eps k^2 {\Dt} \,
\zeta^{(n+1/2)} \right) \, , \\
\label{eq:L1d}
& \frac{L^{(n+1)} - L^{(n)}}{\Dt} = \Mterm^{(n+1/2)} \, , \\
\label{eq:th1d}
& \frac{ \mu_k^{(n+1)} \thhat_k^{(n+1)} - \mu_k^{(n)} \thhat_k^{(n)}}{\Dt} =
\mu_k^{(n+1/2)} \hat{\NLterm}_k^{(n+1/2)} \, .
\end{align}
We must eventually compute the geometry at step $n+1$, and so we insert~\eqref{eq:mu1d} into~\eqref{eq:th1d} and solve for $\theta^{(n+1)}$, giving
\todo[inline]{are we solving for $\theta$ or $\hat{\theta}_k?$.}
\begin{align}
\thhat_k^{(n+1)} =  \thhat_k^{(n)} \exp \left( - \epsilon k^2 \Dt
\elfun^{(n+1/2)} \right) + \Dt \, \hat{\NLterm}_k^{(n+1/2)} \exp \left(
- \frac{1}{2} \eps k^2 \Dt \left( 2 \elfun^{(n+1/2)} - \elfun^{(n)}
\right) \right) \, .
\end{align}
Once again, both exponential terms act to smooth the geometry. Finally, to track the reference point, $\mean{x}$, we simply apply the same RK2 method to~\eqref{eq:xsurf}.

>>>>

At the first stage of RK2 (the half step) we discretize~\eqref{eq:mu}--\eqref{eqn:thetaevo3} as
\begin{align}
\label{eq:mu12d}
& \mu_k^{(n+1/2)} = \mu_k^{(n)} \exp \left( \eps k^2 \frac{\Dt}{2}
\zeta^{(n)} \right) \, , \\
\label{eq:L12d}
& \frac{L^{(n+1/2)} - L^{(n)}}{\Dt/2} = \Mterm^{(n)} \, , \\
\label{eq:th12d}
& \frac{ \mu_k^{(n+1/2)} \thhat_k^{(n+1/2)} - \mu_k^{(n)} \thhat_k^{(n)}}{\Dt/2} 
= \mu_k^{(n)} \hat{\NLterm}_k^{(n)} \, ,
\end{align}
where the superscript indicates the time step. At the next stage of RK2
it will be necessary to evaluate derivatives at the half step $n+1/2$, thus requiring a call to the Stokes solver for which the body shapes must be known. We therefore insert~\eqref{eq:mu12d} into~\eqref{eq:th12d} and solve for $\thhat_k^{(n+1/2)}$, giving
\begin{align}
\label{eq:th12}
\thhat_k^{(n+1/2)} = \left( \thhat_k^{(n)} + \frac{1}{2} \Dt \, \hat{\NLterm}_k^{(n)} \right)
\exp \left( -\epsilon k^2 \frac{\Dt}{2} \elfun^{(n)}\right) \, .
\end{align}
Since $\zeta >0$, the exponential term always {\em suppresses} high-frequencies (exponentially with $k^2$). In other words, this is a smoothing term. With the geometry at step $n+1/2$ known, a call to the Stokes solver allows the computation of $\Mterm^{(n+1/2)}$, $\widehat{\NLterm}_k^{(n+1/2)}$, and $\zeta^{(n+1/2)}$.

At the second stage of RK2, we discretize~\eqref{eq:mu}--\eqref{eqn:thetaevo3} as 
\begin{align}
\label{eq:mu1d}
& \mu_k^{(n+1)} = \mu_k^{(n)} \exp \left( \eps k^2 {\Dt} \,
\zeta^{(n+1/2)} \right) \, , \\
\label{eq:L1d}
& \frac{L^{(n+1)} - L^{(n)}}{\Dt} = \Mterm^{(n+1/2)} \, , \\
\label{eq:th1d}
& \frac{ \mu_k^{(n+1)} \thhat_k^{(n+1)} - \mu_k^{(n)} \thhat_k^{(n)}}{\Dt} =
\mu_k^{(n+1/2)} \hat{\NLterm}_k^{(n+1/2)} \, .
\end{align}
We must eventually compute the geometry at step $n+1$, and so we insert~\eqref{eq:mu1d} into~\eqref{eq:th1d} and solve for $\thhat_k^{(n+1)}$, giving
\begin{align}
\thhat_k^{(n+1)} =  \thhat_k^{(n)} \exp \left( - \epsilon k^2 \Dt
\elfun^{(n+1/2)} \right) + \Dt \, \hat{\NLterm}_k^{(n+1/2)} \exp \left(
- \frac{1}{2} \eps k^2 \Dt \left( 2 \elfun^{(n+1/2)} - \elfun^{(n)}
\right) \right) \, .
\end{align}
Once again, both exponential terms act to smooth the geometry. Finally, to track the reference point, $\mean{x}$, we simply apply the same RK2 method to~\eqref{eq:xsurf}.
